# Products and Deliverables

## R Shiny App
The Outdoor Equity Shiny App has the following functionality:

- Visualize statistical distributions of RIDB and US Census variables
- Visualize relationships between RIDB and US census variables
- Created visualizations can be downloaded as png
- View visitorshed maps of reservable sites both nationally and within state site is located in
- Download customized subsets of the data

Help documentation:

- User guide
- Metadata 
- Background 

## Metadata

```{r, echo=FALSE}
## Note (HD): Should I move this to index.Rmd? ##
# get metadata csv and create csv in r
data_joined_2018_metadata <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/19ZQiaTlSoo1-KlPibeIoVbo0IT0ZU2soFa0XGAEHF88/edit?usp=sharing")

data_joined_2018_metadata_r <- data_joined_2018_metadata %>% 
  write_csv(here("../outdoor-equity/outdoor-equity-app/oe_app/data/data_joined_2018_metadata.csv"))

# read in metadata
data_joined_2018_metadata_df <- read_csv(here("../outdoor-equity/outdoor-equity-app/oe_app/data/data_joined_2018_metadata.csv"))
```


```{r}
# create table of metadata
knitr::kable(data_joined_2018_metadata_df,
             align = "l",
             caption = "Metadata of joined RIDB-ACS dataset",
             row.names = FALSE)
```



