# Products and Deliverables

## R Shiny App
The [Outdoor Equity App](https://shinyapps.bren.ucsb.edu/oe_app/) was built using the `shiny` package [@R-shiny] and has the following functionality:

- Visualize statistical distributions of [RIDB](https://ridb.recreation.gov/download) and [US Census American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/data.html) variables
- Visualize relationships between RIDB and ACS variables
- Download visualizations as PNG images
- View visitorshed maps of reservable sites both nationally and within the state where the campsite is located
- Download customized subsets of the data


## Metadata

```{r, echo=FALSE, message=FALSE}
# read in metadata
data_joined_2018_metadata_df <- 
  read_csv(here("../outdoor-equity/outdoor-equity-app/oe_app/data/data_joined_2018_metadata.csv")) %>% 
  rename("Variable Name" = variable_name,
         "Data Source" = data_source,
         Definition = definition,
         Unit = unit,
         Class = class,
         "Data Type" = data_type,
         Provenance = provenance_filepath,
         "Missing Value Code" = missing_value_code,
         "Total Unique Values" = total_unique_values)
```


```{r, echo=FALSE}
# create table of metadata
knitr::kable(data_joined_2018_metadata_df,
             align = "l",
             caption = "Metadata of joined RIDB-ACS dataset",
             row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("responsive"), full_width = F, position = "left")
```


## Package Versions

```{r, echo=FALSE}
r_packages_df <- 
  tribble(
    ~"R Package", ~Version, ~Purpose,
    "bslib", "0.3.1", "Web Application (theme)",
    "collections", "0.3.5", "Web Application (dictionary)",
    "DT", "0.20", "Web Application (tables)",
    "formattable", "0.2.1", "Web Application (tables)",
    "googlesheets4", "1.0.0", "Web Application (metadata)",
    "here", "1.0.1", "Data cleaning (relative file paths)",
    "janitor", "2.1.0", "Data cleaning (data frame)",
    "lubridate", "1.7.10", "Data cleaning (dates)",
    "plotly", "4.10.0", "Data visualiation (plots)",
    "reactlog", "1.1.0", "Web Application (testing)",
    "rmapshaper", "0.4.5", "Web Application (load time)",
    "rsconnect", "0.8.25", "Web Application (deploy app)",
    "scales", "1.1.1", "Data visualiation (plots)",
    "sf", "1.0.7", "Data cleaning and visualization (maps)",
    "shiny", "1.7.1", "Web Application (build app)",
    "shinycssloaders", "1.0.0", "Web Application (plot loader)",
    "shinydashboard", "0.7.2", "Web Application (dashboard layout elements)",
    "shinydashboardPlus", "2.0.3", "Web Application (box elements)",
    "shinyjs", "2.1.0", "Web Application (hide and show boxes in ui)",
    "shinyWidgets", "0.6.4", "Web Application (inputs)",
    "tidycensus", "1.2.1", "Data acquisition (US census data)",
    "tidyverse", "1.3.1", "Data cleaning, analysis, and visualization",
    "tigris", "1.6", "Data acquisition (state and ZIP code geometries)",
    "tmap", "3.3.2", "Data visualization (maps)",
    "vroom", "1.5.7", "Data cleaning (read in data)",
    "zipcodeR", "0.3.3", "Data acquisition (state info about ZIP code)"
  )

knitr::kable(r_packages_df,
             align = "l",
             caption = "Table includes information on packages used to create the Outdoor Equity App.",
             row.names = FALSE) %>% 
  kable_styling(bootstrap_options = c("responsive"), full_width = F, position = "left")
```

