# Products and Deliverables

## R Shiny App
The [Outdoor Equity App](https://shinyapps.bren.ucsb.edu/oe_app/) was built using the `shiny` package [@R-shiny] and has the following functionality:

- Visualize statistical distributions of [RIDB](https://ridb.recreation.gov/download) and [US Census American Community Survey (ACS)](https://www.census.gov/programs-surveys/acs/data.html) variables
- Visualize relationships between RIDB and ACS variables
- Download visualizations as PNG images
- View visitorshed maps of reservable sites both nationally and within the state where the campsite is located
- Download customized subsets of the data


## Metadata

```{r, echo=FALSE, message=FALSE}
## Note (HD): Should I move this to index.Rmd? ##
# get metadata csv and create csv in r
# data_joined_2018_metadata <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/19ZQiaTlSoo1-KlPibeIoVbo0IT0ZU2soFa0XGAEHF88/edit?usp=sharing")
# 
# data_joined_2018_metadata_r <- data_joined_2018_metadata %>% 
#   write_csv(here("../outdoor-equity/outdoor-equity-app/oe_app/data/data_joined_2018_metadata.csv"))

# read in metadata
data_joined_2018_metadata_df <- 
  read_csv(here("../outdoor-equity/outdoor-equity-app/oe_app/data/data_joined_2018_metadata.csv")) %>% 
  rename("Variable Name" = "variable name",
         "Data Source" = data_source,
         Definition = definition,
         Unit = unit,
         Class = class,
         "Data Type" = data_type,
         Provenance = "provenance (which scripts if used to created this var)",
         "Missing Value Code" = missing_value_code,
         "Total Unique Values" = total_unique_values)
```


```{r, echo=FALSE}
# create table of metadata
knitr::kable(data_joined_2018_metadata_df,
             align = "l",
             caption = "Metadata of joined RIDB-ACS dataset",
             row.names = FALSE) %>%
  kable_styling(latex_options = "scale_down") %>% 
  scroll_box(width = "500px")
```


## Package Versions

The table below includes information on the specific package versions used to create the Outdoor Equity App. 

```{r}

```

